Towards Detecting Stalkerware Through Network Traces

### 2019-03-22 paper/main.tex
Moved filters to appendix

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as "reskinning:" a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven't yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the "ILFMobile group" consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	Spyzie group: Spyzie and SpyMyFone also have identical network signatures 
	as well as identical websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

All of our filters have been collected and are listed here. These can be directly placed in Wireshark. 
\begin{lstlisting}
ip.host contains dedicated.codero.net 
|| http.host contains hellospy
|| (ip.host contains data-api.spymyfone.com && ssl.handshake.extension.data contains "my.spymyfone.com") 
|| (ip.host contains ws.spyzie.com && ssl.handshake.extension.data contains "ws.spyzie.com") 
|| ip.host contains data-api.spymyfone.com 
|| ip.host contains ws.spyzie.com 
|| ip.host contains phonedata.me 
|| ssl.handshake.extension.data contains "phonedata.me" 
|| ip.host contains apiv4.alter757.info 
|| ssl.handshake.extension.data contains "thd.cc" 
|| ip.host contains thd.cc 
|| ip.host contains mspyonline.com 
|| ip.host contains corexchange.com && ssl.handshake.extensions_server_name contains messages02.smstracker.com 
|| ip.host contains life360.helpshift.com 
|| ip.host contains antitheft.ns1.ff.avast.com 
|| (ip.host contains production-finderapps && ssl.handshake.extensions_server_name contains "api.familysafetyproduction.com") || ip.host contains production-finderapps 
|| ssl.handshake.extensions_server_name contains "celulocen.zendesk.com" 
|| ip.host contains www.zoemob.com 
|| ip.host contains unifiedlayer.com 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
|| (ip.dst_host matches "spoc-mu(|w|e)1.norton" && http.user_agent matches "(android|ios)")
|| ssl.handshake.extensions_server_name matches "family.norton" 
|| (udp && ip.host contains stun.l.google.com) 
|| http.host == relay.google.com && http.user_agent matches test 
|| http.host matches "97.logger.mobi" 
|| ssl.handshake.extensions_server_name matches "gpswox" 
|| ip.dst_host matches "gpswox" 
|| ssl.handshake.extensions_server_name matches "geozilla" 
|| ip.dst_host matches "clients.your-server.de" 
|| (ip.addr == 66.147.232.135 && tcp.port == 8080 && http.request.uri contains "signalr") 
|| ip.addr == 66.147.232.135 && tcp.port == 8080 
|| ssl.handshake.extensions_server_name matches "findmykids" 
|| ssl.handshake.extension.data contains "findmykids"  
||  ip.dst_host matches "(trackmyphones.com)" 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
 \end{lstlisting}

@@

For the apps whose traffic do not contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.
\begin{lstlisting}
http.host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| ssl.handshake.extensions_server_name matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| http.user_agent matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| ipv6.dst_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.dst_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"  
|| ipv6.src_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.src_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
\end{lstlisting}

@@

>>>>

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as "reskinning:" a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven't yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the "ILFMobile group" consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	\item \textbf{Spyzie group}: Spyzie and SpyMyFone also have identical network signatures 
	as well as similar websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

@@

For the apps whose traffic do not contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.

@@

\section{Wireshark Display Filters for Specific Signatures}
\begin{lstlisting}
ip.host contains dedicated.codero.net 
|| http.host contains hellospy
|| (ip.host contains data-api.spymyfone.com && ssl.handshake.extension.data 
contains "my.spymyfone.com") 
|| (ip.host contains ws.spyzie.com && ssl.handshake.extension.data contains 
"ws.spyzie.com") 
|| ip.host contains data-api.spymyfone.com 
|| ip.host contains ws.spyzie.com 
|| ip.host contains phonedata.me 
|| ssl.handshake.extension.data contains "phonedata.me" 
|| ip.host contains apiv4.alter757.info 
|| ssl.handshake.extension.data contains "thd.cc" 
|| ip.host contains thd.cc 
|| ip.host contains mspyonline.com 
|| ip.host contains corexchange.com && ssl.handshake.extensions_server_name 
contains messages02.smstracker.com 
|| ip.host contains life360.helpshift.com 
|| ip.host contains antitheft.ns1.ff.avast.com 
|| (ip.host contains production-finderapps && 
ssl.handshake.extensions_server_name contains "api.familysafetyproduction.com") 
|| ip.host contains production-finderapps 
|| ssl.handshake.extensions_server_name contains "celulocen.zendesk.com" 
|| ip.host contains www.zoemob.com 
|| ip.host contains unifiedlayer.com 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
|| (ip.dst_host matches "spoc-mu(|w|e)1.norton" && http.user_agent matches 
"(android|ios)")
|| ssl.handshake.extensions_server_name matches "family.norton" 
|| (udp && ip.host contains stun.l.google.com) 
|| http.host == relay.google.com && http.user_agent matches test 
|| http.host matches "97.logger.mobi" 
|| ssl.handshake.extensions_server_name matches "gpswox" 
|| ip.dst_host matches "gpswox" 
|| ssl.handshake.extensions_server_name matches "geozilla" 
|| ip.dst_host matches "clients.your-server.de" 
|| (ip.addr == 66.147.232.135 && tcp.port == 8080 && http.request.uri contains 
"signalr") 
|| ip.addr == 66.147.232.135 && tcp.port == 8080 
|| ssl.handshake.extensions_server_name matches "findmykids" 
|| ssl.handshake.extension.data contains "findmykids"  
||  ip.dst_host matches "(trackmyphones.com)" 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
\end{lstlisting}

\section{General Signature Wireshark Display Filters}

\begin{lstlisting}
http.host matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"

|| ssl.handshake.extensions_server_name matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"

|| http.user_agent matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"

|| ipv6.dst_host matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.dst_host matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"

|| ipv6.src_host matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.src_host matches 
"(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
\end{lstlisting}
